Active Learning for SAT Solver Benchmarking

### 2022-10-10 paper/paper.tex
section3.2

@@

We transform the runtimes into discrete runtime classes which are determined on a per instance basis.
For each instance $e \in \mathcal{I}$, we use a hierarchial clustering algorithm to assign the runtimes in $\{ r(a, e) \mid a \in A \}$ to one of the $k$ clusters $C_1, \dots, C_k$.
Without loss of generality, the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
For instances with many timeouts, we manually fix the clustering results such that all timeouts $\tau$ are in cluster $C_k$.
The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows.
$$\gamma_k(a, e) = j ~\Leftrightarrow~ r(a, e) \in C_j$$

@@

To determine a ranking based on those labels, we define
\begin{equation}
  \operatorname{score}\!\left(a\right) := \frac{1}{|\mathcal{I}|} \sum_{e \in \mathcal{I}} w_{a,i} \cdot \gamma(a, i)
  \label{eq:rankingeq}
\end{equation}
with $w_{a,i} := 2$ if $\gamma\!\left(a, i\right) = k$ else $w_{a,i} := 1$.
To be applicable to $\hat{a}$, we extend \emph{score} by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.

Discretization would not provide much value if the resulting cluster labels were not predictive of the PAR-2 ranking we seek to predict.
In our work, we use a hierarchical-clustering approach with a single-link criterion for $k = 3$.
For the SAT Competition~2022 Anniversary Track instances~\cite{sat2022}, this ranking approach is capable of correctly deciding for almost all pairs of solvers which one is faster (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$).
Moreover, it decides all solvers correctly if the difference in the cluster-label-ranking scores between two solvers is at least \SI{0.11}{}.
%Refer to Table~\ref{tab:discreteranking} for a comparison with the PAR-2 ranking.
The Spearman correlation~\cite{de2016comparing} of cluster-label and PAR-2 ranking is about \SI{0.988} (in comparison to a perfect value of \SI{1}).
This shows that monotonicity and score distances between ranks align quite well between both ranking methods.

>>>>

To be applicable to $\hat{a}$, we extend \emph{score} by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup 
\left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.

@@

We transform the runtimes into discrete runtime classes which are determined on a per instance basis.
For each instance $e \in \mathcal{I}$, we use a single-link hierarchial clustering algorithm to assign the runtimes in $\{ r(a, e) \mid a \in A \}$ to one of the $k$ clusters $C_1, \dots, C_k$.
Without loss of generality, the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
For instances with many timeouts, we have to manually fix the clustering results such that all timeouts $\tau$ are in cluster $C_k$.
The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows.
$$\gamma_k(a, e) = j ~\Leftrightarrow~ r(a, e) \in C_j$$

@@

To determine solver ranks, we define the adapted scoring function $s : A \rightarrow [1, 2 \cdot k]$ as follows.
\begin{align*}
  \operatorname{s}(a) &:= \frac{1}{|\mathcal{I}|} \sum_{e \in \mathcal{I}} \gamma'_k(a, e)
  &&
  \gamma'_k(a, e) &:= \begin{cases}
  	2 \cdot \gamma_k(a, e) 	& \text{if } \gamma_k(a, e) = k\\
	\gamma_k(a, e)	& \text{otherwise}
  \end{cases}
  \label{eq:rankingeq}
\end{align*}

Discretization would not provide much value if the resulting cluster labels were not predictive of the PAR-2 ranking we seek to predict.
For the SAT Competition~2022 Anniversary Track instances~\cite{sat2022}, this ranking approach is capable of correctly deciding for almost all pairs of solvers which one is faster (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$).
Moreover, it decides all solvers correctly if the difference in the cluster-label-ranking scores between two solvers is at least \SI{0.11}{}.
%Refer to Table~\ref{tab:discreteranking} for a comparison with the PAR-2 ranking.
The Spearman correlation~\cite{de2016comparing} of cluster-label and PAR-2 ranking is about \SI{0.988} (in comparison to a perfect value of \SI{1}).
This shows that monotonicity and score distances between ranks align quite well between both ranking methods.
